Accuracy and precision of zero-heat-flux temperature monitoring

### 2019-11-28 manuscript/manuscript.Rmd
fixed megan's edits

Sixteen studies were included. The primary comparison of zero-heat-flux versus core temperature measurements consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. Data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included. The pooled estimate for the mean bias was `r primary$bias_mean`°C. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C. The GRADE evidence quality rating was downgraded to moderate due to concerns about study limitations. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QADAS-2 were similar to the primary analysis.

@@

Two reviewers independently assessed the risk of bias for the included studies using the revised Quality Assessment of Diagnostic Accuracy Studies (QUADAS-2).[@whiting2011quadas] Reviewers rated the risk of bias for patient selection, conduct of the zero-heat-flux measurements, conduct of the comparator thermometer measurements, and timing and flow (eg, timing of zero-heat-flux and established core temperature measurements, dropouts) as “high”, “low” or “unclear” risk of bias. We worked to minimize the risk of publication bias by conducting a comprehensive search of multiple databases as well as an international clinical trial registry.[@glasziou2001systematic] Statistical approaches for detection of reporting bias were not conducted due to lack of validated methods.[@begg2005systematic] Simulations have revealed that tests for detecting funnel plot asymmetry will result in publication bias being incorrectly identified too often.[@deeks2005performance]

@@

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses[@team2017r] All data and R code used in the meta-analyses can be located [here](https://github.com/awconway/zero-heat-flux-review).

Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator thermometer minus zero-heat-flux thermometer measured in °C. In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'. Other studies reported multiple sets of results, whereby analyses were conducted between zero-heat-flux and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were a part of separate meta-analysis groups. One study reported intraoperative, postoperative and overall results for the same participants. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups.

The conventionally cited clinically acceptable agreement between zero-heat-flux and comparator devices is 0.5 °C.[@eshraghi2014] It was deemed that outer confidence bounds for 95% limits of agreement between zero-heat-flux and core temperature measurements (termed as ‘population limits of agreement’) outside of these bounds would be clinically unacceptable.

@@

A graphical overview of the findings for the primary analysis is presented in Figure \@ref(fig:fig-2). The pooled estimate for the mean bias between zero-heat-flux and core temperature measurements was `r primary$bias_mean`°C. The population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` studies). The mean bias was again `r low_risk$bias_mean`°C with population limits of agreement spanning from `r low_risk$CI_L_rve`°C to `r low_risk$CI_U_rve`°C. A further sensitivity analysis excluding studies that received industry funding revealed population limits of agreement that were considerably wider than the primary analysis (`r conflict$CI_L_rve`°C to `r conflict$CI_U_rve`°C). The GRADE rating for this sensitivity was downgraded to low quality though, due to concerns about study limitations, inconsistency and imprecision.

@@

Table 1 presents results of the primary, sensitivity and all subgroup analyses for the primary comparison as well as secondary comparisons. We conducted two subgroup analyses for the primary comparison according to the clinical setting in which the study was conducted. In the subset of studies conducted in the ICU, the mean bias was `r ICU$bias_mean`°C with population limits of agreement between `r ICU$CI_L_rve`°C and `r ICU$CI_U_rve` °C. The GRADE rating for quality of evidence was downgraded to very low quality due to study limitations, inconsistency and imprecision. In the subset of studies that evaluated the use of zero-heat-flux thermometers during surgery found a mean bias of `r OT$bias_mean`°C with population limits of agreement from `r OT$CI_L_rve`°C to `r OT$CI_U_rve` °C. The quality of evidence rated as moderate quality due to concerns about study limitations.

@@

The mean bias between zero-heat-flux and nasopharyngeal temperature measurements in meta-analysis of results from `r NPA$Studies` studies was `r NPA$bias_mean` °C. Population limits of agreement were `r NPA$CI_L_rve`°C to `r NPA$CI_U_rve`°C. We downgraded the quality of evidence to low, again due to concerns about imprecision and study limitations.

@@

Results from this systematic review have important implications for practice. Clinicians should consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as one degree Celsius higher or lower than core temperature. It was reassuring that results of our sensitivity analysis restricted to studies judged to be at low risk of bias using the QADAS-2 tool were similar. As such, it is vital for clinicians considering using this device to first determine if differences in temperature smaller than this magnitude would be important for the given clinical situation. If so, then it may not be appropriate to substitute the zero-heat-flux device in place of a core thermometer.

Our estimates of the accuracy of zero-heat-flux thermometers are similar to results from a previous meta-analysis that compared other peripheral thermometers, such as sublingual and temporal artery devices, with core temperature measurements.[@niven2015accuracy] However, this previous meta-analysis used a statistical approach that did not incorporate the magnitude of heterogeneity in results between studies or sampling error. As such, it is possible that the zero-heat-flux thermometer may still more precise than other peripheral thermometers.

@@

Although comparisons with core temperature are the most important to evaluate, we included studies that compared zero-heat-flux and peripheral temperature monitoring devices in our review because clinicians may be interested to determine if this new technology is sufficiently accurate to replace other ways to non-invasively monitor temperature. Also, there may be situations where it is not practical to expose research participants to the potential additional risk of invasive core temperature monitoring for the purposes of evaluating the new temperature monitoring device when it is standard practice for only peripheral temperature monitoring to be used. Whether or not the zero-heat-flux thermometer is sufficiently accurate to be used in place of nasopharyngeal thermometers is unclear. Population limits of agreement calculated from four studies were broad, spanning from one degree less than zero-heat-flux temperature to one degree higher. Further studies are therefore required.

@@

The accuracy of zero-heat-flux temperature monitoring has been evaluated in various intraoperative-operative contexts as well as in the intensive care unit setting. Our subgroup analyses did not indicate that the accuracy of this temperature monitoring device was any more effective in a particular setting. Only one study included in our systematic review included pediatric patients. Additional studies to evaluate the accuracy of this device in children may be warranted to increase confidence.

>>>>

Sixteen studies were included. The primary comparison of zero-heat-flux versus core temperature measurements consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. Data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included. The pooled estimate for the mean bias was `r primary$bias_mean`°C. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C. The GRADE evidence quality rating was downgraded to moderate due to concerns about study limitations. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis.

@@

Two reviewers independently assessed the risk of bias for the included studies using the revised Quality Assessment of Diagnostic Accuracy Studies (QUADAS-2).[@whiting2011quadas] Reviewers rated the risk of bias for patient selection, conduct of the zero-heat-flux measurements, conduct of the comparator thermometer measurements, and timing and flow (eg, timing of zero-heat-flux and established core temperature measurements, dropouts) as 'high', 'low' or 'unclear' risk of bias. We worked to minimize the risk of publication bias by conducting a comprehensive search of multiple databases as well as an international clinical trial registry.[@glasziou2001systematic] Statistical approaches for detection of reporting bias were not conducted due to lack of validated methods.[@begg2005systematic] Simulations have revealed that tests for detecting funnel plot asymmetry will result in publication bias being incorrectly identified too often.[@deeks2005performance]

@@

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses.[@team2017r] All data and R code used in the meta-analyses can be located [here](https://github.com/awconway/zero-heat-flux-review).

Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator thermometer minus zero-heat-flux thermometer measured in degrees Celsius (°C). In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'. Other studies reported multiple sets of results, whereby analyses were conducted between zero-heat-flux and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were a part of separate meta-analysis groups. One study reported intraoperative, postoperative and overall results for the same participants. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups.

The conventionally cited clinically acceptable agreement between zero-heat-flux and comparator devices is 0.5°C.[@eshraghi2014] It was deemed that outer confidence bounds for 95% limits of agreement between zero-heat-flux and core temperature measurements (termed as ‘population limits of agreement’) outside of these bounds would be clinically unacceptable.

@@

A graphical overview of the findings for the primary analysis is presented in Figure \@ref(fig:fig-2). The pooled estimate for the mean bias between zero-heat-flux and core temperature measurements was `r primary$bias_mean`°C. The population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` studies). The mean bias was again `r low_risk$bias_mean`°C with population limits of agreement spanning from `r low_risk$CI_L_rve`°C to `r low_risk$CI_U_rve`°C. A further sensitivity analysis excluding studies that received industry funding revealed population limits of agreement that were considerably wider than the primary analysis (`r conflict$CI_L_rve`°C to `r conflict$CI_U_rve`°C). The GRADE rating for this sensitivity was downgraded to low quality though, due to concerns about study limitations, inconsistency and imprecision.

@@

Table 1 presents results of the primary, sensitivity and all subgroup analyses for the primary comparison as well as secondary comparisons. We conducted two subgroup analyses for the primary comparison according to the clinical setting in which the study was conducted. In the subset of studies conducted in the ICU, the mean bias was `r ICU$bias_mean`°C with population limits of agreement between `r ICU$CI_L_rve`°C and `r ICU$CI_U_rve` °C. The GRADE rating for quality of evidence was downgraded to very low quality due to study limitations, inconsistency and imprecision. In the subset of studies that evaluated the use of zero-heat-flux thermometers during surgery found a mean bias of `r OT$bias_mean`°C with population limits of agreement from `r OT$CI_L_rve`°C to `r OT$CI_U_rve`°C. The quality of evidence was rated as moderate quality due to concerns about study limitations.

@@

The mean bias between zero-heat-flux and nasopharyngeal temperature measurements in meta-analysis of results from `r NPA$Studies` studies was `r NPA$bias_mean`°C. Population limits of agreement were `r NPA$CI_L_rve`°C to `r NPA$CI_U_rve`°C. We downgraded the quality of evidence to low, again due to concerns about imprecision and study limitations.

@@

Results from this systematic review have important implications for practice. Clinicians should consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as 1°C higher or lower than core temperature. It was reassuring that results of our sensitivity analysis restricted to studies judged to be at low risk of bias using the QUADAS-2 tool were similar. As such, it is vital for clinicians considering using this device to first determine if differences in temperature smaller than this magnitude would be important for the given clinical situation. If so, then it may not be appropriate to substitute the zero-heat-flux device in place of a core thermometer.

Our estimates of the accuracy of zero-heat-flux thermometers are similar to results from a previous meta-analysis that compared other peripheral thermometers, such as sublingual and temporal artery devices, with core temperature measurements.[@niven2015accuracy] However, this previous meta-analysis used a statistical approach that did not incorporate the magnitude of heterogeneity in results between studies or sampling error. As such, it is possible that the zero-heat-flux thermometer may still be more precise than other peripheral thermometers.

@@

Although comparisons with core temperature are the most important to evaluate, we included studies that compared zero-heat-flux and peripheral temperature monitoring devices in our review because clinicians may be interested to determine if this new technology is sufficiently accurate to replace other ways to non-invasively monitor temperature. Also, there may be situations where it is standard practice for only peripheral temperature monitoring to be used. In these circumstances, it is not practical to expose research participants to the potential risks of invasive core temperature monitoring for the purpose of evaluating the new temperature monitoring device. Whether or not the zero-heat-flux thermometer is sufficiently accurate to be used in place of nasopharyngeal thermometers is unclear. Population limits of agreement calculated from four studies were broad, spanning from 1°C less than zero-heat-flux temperature to one degree higher. Further studies are therefore required.

@@

The accuracy of zero-heat-flux temperature monitoring has been evaluated in various intraoperative contexts as well as in the intensive care unit setting. Our subgroup analyses did not indicate that the accuracy of this temperature monitoring device was any more effective in a particular setting. Only one study included in our systematic review included pediatric patients. Additional studies to evaluate the accuracy of this device in children may be warranted to increase confidence.
